Text embedding models yield high-resolution insights into conceptual knowledge from short multiple-choice quizzes

### 2024-02-21 paper/main.tex
progress on WIP GLMM methods section

As necessary, we then iteratively removed random effects from the maximal model until it successfully converged with a full rank (i.e., non-singular) random effects variance-covariance matrix. 
%When inspecting the model's random effect estimates revealed multiple terms estimated at the boundary of their parameter space (i.e., variance components of 0 or correlation terms of  $\pm 1$), we found that the order in which we eliminated these terms typically did not affect which terms did and did not need to be removed in order for the model to converge to a non-degenerate solution.
When this required eliminating multiple terms whose estimates reached the boundary of their parameter space (i.e., variance components of 0 or correlation terms of  $\pm 1$), we found (qualitatively) that the order in which we did so typically did not affect the set of terms that needed to be removed in order for the model to converge to a non-degenerate solution.
However, in order to ensure consistency in our approach across the 15 separate GLMMs we fit in this set of analyses, we established a general approach to selecting a boundary-estimated parameter to eliminate for each stepwise reduction of the model's complexity, broadly adapted from~\citet{BateEtal15b}.
First, we constrained any correlation terms estimated at the boundary (i.e., $\pm 1$) to 0 and re-fit the model.
This typically resulted in the variance component for the slope or intercept (or possibly both) associated with the zero-constrained correlation being estimated as 0, suggesting which terms we should consider removing.
Second, when choosing which of two variance components to remove from the model, we prioritized removing higher-order terms before removing lower-order terms.
In other words, if the variance components for a random slope and a random intercept were both estimated as 0, we chose to remove the random slope.
Additionally, if constraining a correlation term to 0 allowed the model to converge with a full rank variance-covariance matrix, we left the correlation term constrained rather than removing the associated slope and/or intercept component.
Third, to choose which of two zero-estimated variance components of the \textit{same} order to remove from the model, we took one of two approaches.
If the two components were both random slopes, performed a Principal Components Analysis (PCA) on the random effects variance-covariance matrix, and compared the proportion of variance explained by the second component from each of the two random effect groupings. 
If one If one explains a substantially smaller proportion of variance than the other, we dropped its corresponding random slope.
Alternatively, if the difference in variance explained by the two is (qualitatively) small, or if the two variance components we wanted to choose between were random intercepts rather than slopes, we removed the one whose absence resulted in the greater decrease in the Akaike information criterion (AIC).

>>>>

When necessary, we then iteratively removed random effects from the maximal model until it successfully converged with a full rank (i.e., non-singular) random effects variance-covariance matrix. 
%When inspecting the model's random effect estimates revealed multiple terms estimated at the boundary of their parameter space (i.e., variance components of 0 or correlation terms of  $\pm 1$), we found that the order in which we eliminated these terms typically did not affect which terms did and did not need to be removed in order for the model to converge to a non-degenerate solution.
When this required eliminating multiple terms whose estimates reached the boundary of their parameter space (i.e., variance components of 0 or correlation terms of  $\pm 1$), we found (qualitatively) that the order in which we did so typically did not change the set of terms that needed to be removed in order for the model to converge to a non-degenerate solution.
However, in order to ensure we fit all of the models for these analyses in a consistent way, we devised some simple heuristics for choosing which boundary-estimated parameter to eliminate in each stepwise reduction of the model's complexity, based on a combination of popular approaches~\citep{BarrEtal13,BateEtal15b,MatuEtal17}.
First, we constrained any correlation terms estimated at their boundaries (i.e., $\pm 1$) to 0 and re-fit the model.
This usually resulted in the variance component for the associated random slope (or occasionally the intercept, or even both) being estimated as 0, and suggested which terms we should consider removing.
Second, when choosing which of two zero-variance terms to remove from the model, we prioritized removing higher-order terms before lower-order terms.
In other words, if the estimated variance for both a random intercept and a random slope were 0, we chose to remove the random slope term before re-fitting the model and evaluating whether or not to remove the intercept.
%Additionally, if constraining a boundary-estimated correlation term to 0 allowed the model to converge with a full rank covariance matrix, we left the correlation term constrained rather than removing the associated slope and/or intercept component.
Third, when choosing which of two boundary-estimated variance components of the \textit{same} order to exclude from the model first, we took one of two approaches.
If the two components were random slopes with small but non-zero coefficients (e.g., on the order of $10^{-4}$ or less), we performed a Principal Components Analysis (PCA) on the random effects variance-covariance matrix and compared the variance explained by the two terms' corresponding PCs. 
If one explained a substantially smaller proportion of variance than the other (loosely defined), we dropped its corresponding random slope.
Alternatively, if both slopes' principal components explained a similar amount of variance, or if the two terms we wanted to compare were random intercepts rather than slopes, we removed the one whose exclusion resulted in the greater decrease in the Akaike information criterion (AIC).
Fourth, after removing either a slope or intercept from one random effect grouping, we re-introduced any correlation terms we had previously constrained to 0 for other random effect groupings to check if they could now be estimated. 
Finally, once the model successfully converged with a full rank covariance matrix, we did not remove random effects for any other reasons (e.g., to lowered AIC, increase log-likelihood, etc.). 
We chose this stopping criterion as a conceptual ``middle ground'' between two popular but opposing approaches to model selection that advocate for either retaining the maximal model that allows convergence, regardless of singular fits~\citep[at the potential cost of decreased power; e.g.,~][]{BarrEtal13} or performing additional hypothesis tests on individual parameters and removing non-significant components to achieve a parsimonious model~\citep[at the potential cost of increased Type I error rates; e.g.,~][]{BateEtal15b}.
%We chose this stopping criterion as a conceptual ``middle ground'' between two popular but opposing approaches to model selection that advocate (respectively) for either retaining the maximal model that allows convergence, regardless of singular fits~\citep[at the potential cost of decreased power; e.g.,~][]{BarrEtal13} or testing individual parameters achieving a parsimonious model by discarding all parameters that don't significantly decrease goodness of fit ~\citep[at the potential cost of increased Type I error rates; e.g.,~][]{BateEtal15b}.
Our threshold for inclusion of random effects is intended to achieve a reasonable balance between these trade-offs.

To assess the predictive value of our knowledge estimates for individual quiz questions, we used the same sets of observations used to fit each GLMM to fit a second set of ``null'' models (similarly with a logistic link function). We fit these models with the formula:
\[
    \mathtt{accuracy \sim (1\ \vert\ participant) + (1\ \vert\ question)}
\]

%In order to assess the predictive value of the knowledge estimates, we then fit a second set of ``null'' models to the same sets of observations used to fit our full GLMMs. These 
%
%
%used the same sets of observations used to fit these GLMMs to fit a second set of ``null'' models. 
%
%
%
%Next, in order to assess the predictive value of the knowledge estimates 
%
%
%
%In order to assess the predictive value of the knowledge estimates we used to fit each GLMM, we then fit a \textit{second} model to the same data as each of the 15 
%
%In order to assess whether the knowledge estimates we used to fit each GLMM could reliably predict participants' success on held-out questions, we then fit a second GLMM to the observations
